### 2022-02-04 src/document.tex

% Book information
\titlehead{}
\title{Les besoins en éducation pour les pratiques opensource}
\subtitle{Mémoire de M2 Sciences de l'Éducation, parcours SYNVA}
\author[PH]{Paul Hervot}
\date{\today}
\publishers{Université de Strasbourg} % TODO: is it really the publisher?
% TODO: add INSPÉ and UNISTRA logos to the title page
% TODO: mention the jury on the title page

% Pre-document content, uses roman numerals
\frontmatter

\input{document_start}

% Main document content, resets page numbering and uses arabic numbers
\mainmatter
\setchapterstyle{kao}

\addpart{Partie théorique}

\include{introduction}
\include{literature_review}

\addpart{Notre étude}

\include{methodology}
\include{results}
\include{discussion}

\addpart{Conclusion}

\include{conclusion}

% Number chapters with letters from now on
\appendix

\addpart{Annexes}

\include{annexes}

% End of the main document content
\backmatter

\include{document_end}