### 2022-02-04 src/document_end.tex

\setchapterstyle{plain}

\defbibnote{bibnote}{Références utilisées, présentées par ordre de première citation.\par\bigskip}
\printbibliography[heading=bibintoc, prenote=bibnote]

\printindex

\printglossary