### 2022-02-04 src/glossary.tex

\newacronym{gpl}{GPL}{GNU General Public License}
\newacronym{cc}{CC}{Creative Commons}