### 2022-08-01 src/annexes.tex
appendices: include the data collection code

\setchapterstyle{plain}
\pagelayout{wide}
\chapter{Code d'extraction des données}

\inputminted[fontsize=\footnotesize,linenos,breaklines]{java}{experiment/data_collection/Experiment.java}

\chapter{Code de calcule des résultats}