### 2022-08-18 src/document.tex
manuscript: do a bit of cleaning and TODO-resolving

% Book information
\titlehead{}
\title{L'identification des projets de logiciel libre accessibles aux nouveaux contributeurs}
\subtitle{Mémoire de M2 Sciences de l'Éducation, parcours SYNVA}
\author[PH]{Paul Hervot}
\date{\today}
\publishers{
    \begin{figure}[ht]
        \includegraphics[width=0.5\textwidth]{unistra}
    \end{figure}
}
\dedication{%
    Un grand merci à Antoine Pietri pour son aide précieuse lors de la collecte des données sur le graphe de
    Software Heritage.%
}
% TODO: add INSPÉ and UNISTRA logos to the title page
% TODO: mention the jury on the title page

% Pre-document content, uses roman numerals
\frontmatter

\input{document_start}

% Main document content, resets page numbering and uses arabic numbers
\mainmatter
\setchapterstyle{kao}

\include{introduction}

\addpart{Partie théorique}

\include{literature_review}

\addpart{Partie expérimentale}

\include{methodology}

\include{results}

\include{discussion}

\addpart{Conclusion}

\include{conclusion}

% Number chapters with letters from now on
\appendix

\addpart{Annexes}

\include{annexes}

% End of the main document content
\backmatter

\include{document_end}