### 2022-08-23 main.tex
results: almost finish the results and discussion part

% Suppress some compilation warnings
\RequirePackage[save,showerrors]{silence}
\WarningFilter{biblatex}
  {The starred command '\DeclareDelimAlias*' is} % APA package is using this deprecated starred version
\WarningFilter{transparent}
  {Loading aborted} % I can't find where it is loaded, probably somewhere in kao's cls or sty
\WarningFilter{latex}
  {Unused global option} % They are, but kaobook is silently passing them to its base class
\WarningFilter{latex}
  {Marginpar on page} % Margin stuff moved to not overlap with other stuff, this is fine
\WarningFilter{latex}
  {There were undefined references} % seems spurious, no undefined references are reported
\WarningFilter{biblatex}
  {Please (re)run Biber on the file} % linked to the previous warning

\documentclass[
    a4paper,
    chapterentrydots=true,
    fontsize=11pt,
    open=any,      % Force new chapters to start on any page, not only on right (odd) pages
    secnumdepth=1, % How deep to number headings (1 = sections)
    twoside=true,  % Use different layouts for even and odd pages
]{kaobook}

% TODO: show the page number on each page (unless forbidden by the dissertation requirements)

% Code highlighting
\usepackage{minted}

% Main and default language is french but there are some parts in english
\usepackage[main=french,english]{babel}
\newcommand{\en}[1]{\foreignlanguage{english}{\itshape #1}}
\usepackage{csquotes}
\MakeOuterQuote{"}

% Allow to include SVG with \includesvg similarly to \includegraphics for other image formats
\usepackage[inkscapepath=./build/svg-inkscape/]{svg}

% Bibliography configuration
\usepackage[style=apa]{kaobiblio}
\DefineBibliographyExtras{french}{\restorecommand\mkbibnamefamily}
\addbibresource{references.bib}
% Properly cite articles indirectly cited via meta-analysis, this mimics the \nocitemeta command from apacite
\DeclareBibliographyCategory{meta}
\renewbibmacro*{begentry}{\ifcategory{meta}{\ensuremath{{}^\ast}}{}}
\newcommand*{\nocitemeta}[1]{\nocite{#1}\addtocategory{meta}{#1}}

% Hyperref integration
\usepackage{kaorefs}
\usepackage[nobiblatex]{xurl}

% For typesetting hypothesis
\usepackage[framed=true, hypothesisbackground=gray!20!white]{kaotheorems}
\usepackage{thmtools}
\usepackage{thm-restate}

% For file inclusion and figure formatting
\usepackage{svg} % for svg images
\usepackage{caption}
\usepackage{subcaption}
\usepackage{pdfpages}

% Search paths for images and tex input files
\graphicspath{{images/}}

% Configure the index and the glossary
\makeindex[columns=3, title=Alphabetical Index, intoc]
\makeglossaries
\setglossarystyle{altlist} % TODO: switch to altlistgroup if I end up with a lot entries
\input{glossary}
% shortcut for english glossary entries
\newcommand{\engl}[1]{\en{\gls{#1}}}
\newcommand{\enGl}[1]{\en{\Gls{#1}}}
\newcommand{\englpl}[1]{\en{\glspl{#1}}}
\newcommand{\enGlpl}[1]{\en{\Glspl{#1}}}

% Hyphenation
\hyphenation{Git-Hub}

\begin{document}
    \input{document}
\end{document}